---
title: "Cutpoint Modeling"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Drafting 

there is no clear demarcation in the BMRQ distribution corresponding to anhedonia, meaning that sensitivity to musical reward is a continuum in the population.”
(From perception to pleasure. Zatorre. 2023).



- [Mas Her 2014](https://www.cell.com/current-biology/fulltext/S0960-9822(14)00133-X?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS096098221400133X%3Fshowall%3Dtrue#supplementaryMaterial)
- [Mas Her 2018](https://pdf.sciencedirectassets.com/273019/1-s2.0-S0079612318X00028/1-s2.0-S0079612318300177/main.pdf?X-Amz-Security-Token=IQoJb3JpZ2luX2VjECsaCXVzLWVhc3QtMSJHMEUCIQD20RybUtBp84V1fHZyccuxQRWOGJSNaEVSPxqWyRlh5QIgevfE2mhKxhIPjTstaaHRsm0qtDX9XlgFeuxEqn1TYrYquwUI1P%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FARAFGgwwNTkwMDM1NDY4NjUiDAYbKQw089wvTe6P0SqPBaFuEUfixwIORZI5XugUyl9duLWUsPzofkoXpZLpqKusA1lWIob%2FSbCoKM2XUuUostZS2i1wJkPNmyw%2FttDtwZoeGcZlf6NQWqr%2BZ%2FLdTz8VbhIlH0O6dGbOZSZ%2BArLJLTzF5dh2WILYg2uRrnmJhgYTfju7xKp%2F%2Bd0QfShqcQs6%2FFySq6lBQjq%2BNM0Lfu4bNcGyoX8GJHZ5rjpdJVM7rKDdNNnTLNC4IBNg%2FSZdJSDvcPL5wpWser73Nla8FPCILca7vpnlLaam2bvkcN2kyqtta29b0nceYb5CwHRwO3J5EKuRVLzpNUDoRCedelrOZZj%2FnuVoQsTMJPXGf3ejDY4WMtCasKlImxWWkfORB9u4Jl3vHxxbIR3MwWWo7O84cvyOhoZxMsqG2ljCYCqtht9Y7IyXdi0jbcRdmQ6H0qzYs%2BD9FaIdDEDmxPaMUVdPLEBQ59y0hNNW0YxShQB38Btnw9%2BPiRQaywOKqJFHC5aFTJl8aBCC4m6TA5FFSDeQtjRoA6KXH2LrWH3b6A8I63rBIwtuNzXs%2Fz%2FMyIRhI70caENIo6rkyfrLc8UEg52m6EszKpxpRlC7AbwDzpfOUJvN9e0cZbTBGmMjyd4O9Qkdk3%2Fy47VGQd%2Fb7%2FdRE%2F7DaQ717p%2F32VodMUwMyoD86xeHgPFkmTzq2bkz0W8ANy%2FxnoSi20sQ7N6iMhR51an9R30guXrvT%2BVBDq4sD%2FQ1evb0QJUHO6BNRkLBXO8HjXHjwYb9KJwgEG379%2BkAMs0klNddpF%2FsPskYMcEFqpCZtDn79mAeZB97LTcdBYO2HoKyiFFnYogoca5NALGReaWIyVDjHdrZlMtghFRmnkmvBDTQxkTSRQ%2BP%2FtfccpDMO8YwgPmjrQY6sQHHMljL4WzwtJ%2B8TtTGH%2B2WZvPGNZd2T3xWdj%2BsPBq9laXcghPo46WEbSDg9tc60vQbUWoL3iRH0NRjz0%2FPPMi87lzqWTpohMQ1IGaL6AIkMrenP1Bnrp%2BOzVmzvSdzKr3prr9M0%2BMrbZO1RGZ%2BehhwxysQq1RvgbjUzQOcoKsdmp%2B%2B0VagVu6lAJYasbN%2BgPDcdSW8k%2FV7mf%2FJXkavvfdq%2FiUJenZuJxqWl%2BOZSLgzclM%3D&X-Amz-Algorithm=AWS4-HMAC-SHA256&X-Amz-Date=20240118T111727Z&X-Amz-SignedHeaders=host&X-Amz-Expires=300&X-Amz-Credential=ASIAQ3PHCVTY3VXNKTYK%2F20240118%2Fus-east-1%2Fs3%2Faws4_request&X-Amz-Signature=56f3250da4775aadf6d5e48320a8dfff8407d0d21f02ad2079a2af5e55855129&hash=5de7ce1a4a804e20022a5d3c3a8d35f12141ac9f86b0a91768a4d161cfb7edee&host=68042c943591013ac2b2430a89b270f6af2c76d8dfd086a07176afe7c76c2c61&pii=S0079612318300177&tid=spdf-23b566e9-1de4-42a6-a9e3-a3d8c631e3b3&sid=ca4de142607ec84a91-9ed8-e8094b8d2c55gxrqb&type=client&tsoh=d3d3LnNjaWVuY2VkaXJlY3QuY29t&ua=05015d555157500b5b&rr=84766e806a314188&cc=gb)




- musical anhedonia and hyperdonia : the need for psychometric measurment , construct or statistical artefact? 

- musical anhedonia is condition where individuals do not find music pleasurable 
- perception abilities are fine, but blunted for music specific 
- mas-herro musical cell first do this 
- mas-herro make this claim where they do and find 
- kaithos also follow up on this where futher claim showing should see effect only with pleasing, not displeasing 
- (louis and belfi paper as well)
- in all three papers defined by use BMRQ and PAS to define this with other instruments 
- all papers report descriptives, but dont have lack of correlation between two
- [contact authors, able to get correlcaitons between the two?]
- causal model is that two tools CAUSE A, which then leads to specific effects on the DV 
- A is used as IV, with or without others? 
- Problem is that A has information that is unable to be disentangled from either PAS or eBMRQ unless added
- would not even make sense to include PAS or eBMRQ in model, but would be better modeling choice bc then can estimate where effect is coming from 
- esp if consider degree of correlation in the data orginally 
- main points are
- theoretically good that maybe they are not correlated, but they should be? 



## Process

1. State a clear question
2. Sketch your casual assumptions
3. Use the sketch to define a generative model
4. Use the generative model to build and estimator 


```{r}
library(ggdag)
library(dagitty)
library(ggplot2)
library(cmdstanr)
library(rethinking)
library(bayesplot)
theme_set(theme_dag())
```

### Scientific Question

What does posterior look like if you stratify your IV using two relative cutpoints, then regress onto condition where there is true effect?
Describe the association between adult weight and height using a linear model

### State Causal Assumptions 

- Variables: DV- Plesantness rating, IV- anhedonianess, IV, type of stimuli (please displease)
- know that total eBMRQ squore and PAS score without auditory items are correlated 

- P: DV, pleasentness 
- A: IV, anhedonianess (unobserved, continuous, could be binarized, try for hyperhedonized)
- S: IV, stimuli type (displeasing or displeasing)
- AS: IV, Anhedonia no audio (observed)
- MR: IV, Musical reward (observed)

```{r}

dag <- dagitty("dag{A -> P <- S; U -> P; MR -> A; AS -> A}")

ggdag(dag, layout = "circle") +
  labs(title = "Causual Structure for Anhedonia Modeling") 
  
```

### Sketch Causal Model 

```{r}
library(tidyverse)
# dont model with normal!  Let's just assume normal for now 

# use gamma distribution since will skew positive, assume most people are near clump, with big tail ??

n_samples <- 10000

k_cor2 <- read.csv("data/kaithos_2024.csv", header = FALSE)

names(k_cor2) <- c("PAS", "eBMRQ")

# empirically reported values 
eBMRQ_mean <- 91.32
eBMRQ_sd <- 21.84
PAS_mean <- 9.86
PAS_sd <- 2.53
eBMRQ_PAS_cc <- -0.435

covariance <- eBMRQ_PAS_cc * eBMRQ_sd * PAS_sd

# plot empirical distributions if standrdized

covariance

Sigma <- matrix(c(eBMRQ_sd^2, covariance, covariance, PAS_sd^2), nrow = 2)

xy <- MASS::mvrnorm(n = n_samples, mu = c(eBMRQ_mean, PAS_mean), Sigma = Sigma)

colnames(xy) <- c("PAS", "eBMRQ")

kath_dat <- data.frame(xy) |> as_tibble()

kath_dat

hist(k_cor2$PAS)
hist(k_cor2$eBMRQ)

hist(kath_dat$PAS)
hist(kath_dat$eBMRQ)

# have two varaibles with apprx relationship, but distributions are not totally understood given normal and bounds 

# bascially have some sort of moderation effect 
# try it first just with simple model where can see differences in P 


# dont forget noise 


with(kath_dat, plot(PASz, eBMRQz))

kath_dat$unobserved <- rnorm(n_samples, mean = 0, sd = 3)


# first figure out composite 

# anhedonics - less than on sd on PAS, below tenth percentile on BRMQ

pas_cutpoint <- -.5
bmrq_cutpoint  <- -.5

kath_dat$PASz <- (kath_dat$PAS - mean(kath_dat$PAS)) / sd(kath_dat$PAS)

kath_dat$eBMRQz <- (kath_dat$eBMRQ - mean(kath_dat$eBMRQ)) / sd(kath_dat$eBMRQ)

kath_dat |>
  summarise(min_val = min(PASz))


kath_dat <- kath_dat |>
  mutate(anhed = case_when(
    PASz < pas_cutpoint & eBMRQz < bmrq_cutpoint ~ "Anhedonic",
    .default = "normal"
  ))


kath_dat |>
  ggplot(aes(x = PASz, y = eBMRQz, color = anhed)) +
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm")

kath_dat |>
  ggplot(aes(x = PASz, y = eBMRQz, color = anhed)) +
  geom_point() +
  theme_minimal() +
  geom_smooth(method = "lm")


# Created varaible of anhedonia without tool creates artificial mediator variable that knows everything about upstread 

# if add musical anhedonia varible to model, doesnt even make sense since eBMRQ and PAS have all informaiton, why not model it that way. 

# once learn about Z (artifical cut points), dont need anything more about X (measuremnts)

### stratify by mediator if ONLY intersted in direct effect of plant 

# if stratify by treatement after experiment, dont put it in
# consequences of treatement itself should not be in estimator!!!! 

# stratify by consequence of treatment -- fungus in this case 

# no common cause? 

# A is caused by eBMRQ and PAS 
# X and Y are not related because they don't share a common cause ! (in this case, trying to argue that musical anhedonia is distinct )

# crazy 

causal_effect_ples <- 0.2
kath_dat$pleas_anhed <- causal_effect_ples * kath_dat$PASz + kath_dat$unobserved

model_1 <- lm(pleas_anhed ~ anhed, data = kath_dat)

summary(model_1)

# everything that is known about A come from eBMRQ and PAS?? 



# check each separate, then as composite 

# pas 
kath_dat$pleas_pas <- causal_effect_ples * kath_dat$PASz + kath_dat$unobserved
# bMRQ
kath_dat$pleas_bmrq <- causal_effect_ples * kath_dat$eBMRQ + kath_dat$unobserved





# just general effect of pleasentness with pleasing stimuli 

















eBMRQ <- rnorm(n_samples, mean = 91.32, sd = 21.84)
PAS <- rnorm(n_samples, mean = 9.86, sd = 2.53)


kath_dat







eBMRQ_g <- rgamma(n_samples, shape = 5, rate = .1) + 90 

mean(eBMRQ_g)
sd(eBMRQ_g)



PAS_g <- rgamma(n_samples, mean = 9.86, sd = 2.53)

  


plot(xydf$PAS, xydf$eBMRQ)



mean(eBMRQ)




```















