---
title: "Height Weight Model"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

- next step: actually go through chapter for this 


## Process

1. State a clear question
2. Sketch your casual assumptions
3. Use the sketch to define a generative model
4. Use the generative model to build and estimator 

```{r}
library(rethinking)
data(Howell1)
```

### Scientific Question

Describe the association between adult weight and height. 

```{r}
d <- Howell1[Howell1$age>= 18,]
```


### State Causal Assumptions 

```{r}
library(ggdag)
library(dagitty)
library(ggplot2)
theme_set(theme_dag())

dag <- dagitty("dag{H -> W <- U}")

ggdag(dag, layout = "circle") +
  labs(title = "Height causes Weight + Unobserved Error")

```

### Make Generative Model 

```{r}

sim_weight <- function(H, b, sd){
  U <- rnorm(length(H), 0, sd)
  W <- b*H + U
  return(W)
}
H <- runif(200, min = 130, max = 170)
W <- sim_weight(H, b = 0.5, sd = 5)
plot(W ~ H, col = 2, lwd = 3)

```

### Approxmiate

```{r}
m3.1 <- quap(
  alist(
    W ~ dnorm(mu,sigma),
    mu <- a + b*H,
    a ~ dnorm(0,10),
    b ~ dunif(0,1),
    sigma ~ dunif(0,10)
  ) , data=list(W=W, H=H)
  )

n <- 1e3
a <- rnorm(n, 0, 10)
b <- runif(n, 0, 1)
plot(NULL, xlim=c(130,170), ylim = c(50,90),
                xlab = "height", ylab = "weight")
for (j in 1:50) abline(a = a[j], b=b[j], lwd = 2, col = 2)

precis(m3.1)

post <- extract.samples(m3.1)

head(post)

rethinking::stancode(m3.1)

```

### Re-Build with Stan 

From [Stan User Guide](https://mc-stan.org/docs/stan-users-guide/linear-regression.html) 

```{r}
library(cmdstanr)

m3.1_model <- 
   cmdstan_model(
  "templates/rethinking/cont_IV_DV.stan",
    cpp_options = list(stan_threads = TRUE)
  )

kung_list = list(
  N = length(W),
  w=W,
  h=H)

m3.1_stan <-
  m3.1_model$sample(
    data =
      list(
        N = length(W),
        w = W, 
        h = H
      ),
    chains = 4,
    parallel_chains = 4,
    threads_per_chain = 2,
    ### Slow and reliable
    # adapt_delta = 0.95,
    iter_warmup = 1000,
    iter_sampling = 1000,
    refresh = 50
    ### (Relatively) quick and dirty
    ##iter_warmup = 500,
    ##iter_sampling = 500,
    ## refresh = 25
  )



```


### Rethinking checks but with stan 

```{r}
m3.1

m3.1_stan



model_1_fit$summary(c("variant_difficulties", "person_abilities", "item_difficulties","location_severities")) |>
  select(variable, mean, median, q5,q95, rhat) |>
  print(n = Inf)


```



